Management induced changes of soil organic carbon on global croplands

### 2020-12-14 SOCbudgetPaper/SOCbudget.tex
current status local

In Sect. \ref{sec:carbonbudget} we introduce the basic concept of SOC dynamics as applied in this study and explained in more detail within the refinement of the IPCC guidelines vol.~4 (\citep{calvo_buendia_ipcc_2019}). We additionally describe how we configured and extended the steady-state method (for model code see \citet{mrsoil_2020}). In Sect. \ref{sec:tier1} we shortly refer to the concept of stock change factors as outlined in the Tier 1 approach of the IPCC guidelines (\citep{eggleston_hs_ipcc_2006}, \citep{calvo_buendia_ipcc_2019}).
Section \ref{sec:agrimanagement} provides a detailed description of the global, gridded management data used to drive the model, including crop production levels, residue input rates, manure amendments, and the adoption of irrigation and tillage practices (for model code see \citet{mrcommons_2020}). In Sect. \ref{sec:scenarios} we define the scenarios used to complement our historic model results.

@@

  \begin{table*}[h]
  \caption{Type and data sources for carbon inputs and parameterization to different land-use types }
  \begin{tabular}{p{0.15\textwidth} p{0.20\textwidth} p{0.25\textwidth} p{0.25\textwidth}}
  \tophline
  \textbf{land-use types}   & \textbf{source of carbon inputs} & \textbf{data source} & \textbf{nitrogen and lignin content} \\
  \middlehline
  \multirow{3}{*}{Cropland} & above ground residues & \multirow{3}{*}{\begin{minipage}[t]{0.28\columnwidth}\raggedright
                                                                      \cite{faostat_faostat_2016}, \\
                                                                      \cite{schaphoff_lpjml4_2018} (LPJmL),\\
                                                                      \cite{weindl_livestock_2017} \end{minipage}} & 
                                                      \multirow{2}{*}{\begin{minipage}[t]{0.28\columnwidth}\raggedright
                                                                      lg/c generic values according to Table 5.5B\\
                                                                      in \cite{calvo_buendia_ipcc_2019},\\
                                                                      crop-type specific n/c-ratios from
                                                                      \cite{dietrich_mrcommons_2020}\end{minipage}} \\ \cline{2-2}
                            & below ground residues &  &  \\ \cline{2-2}
                            & manure                &  & Table 5.5C in \cite{calvo_buendia_ipcc_2019} \\ 
                            \hline
  Natural vegetation        & annual litterfall &  \cite{schaphoff_lpjml4_2018} (LPJmL) 
                            & \begin{minipage}[t]{0.28\columnwidth}\raggedright\strut 
                                \cite{calvo_buendia_ipcc_2019} (CENTURY parameterization) \strut \end{minipage}\tabularnewline
 \bottomhline
 \end{tabular}
 \label{tab:datasourceinputs}
 \belowtable{}
 \end{table*}

@@

Note that the absolute changes \(\Delta SOC\) can be also interpreted as the SOC debt (\citep{sanderman_soil_2017}) due to human cropping activities; whereas relative changes \(F^{\mathrm{SCF}}\) can be considered stock change factors as defined within the IPCC guidelines of 2006 \citep{eggleston_hs_ipcc_2006}. Note that the SOC debt is equivalent to the negated cumulative SOC emission (\citep{pugh_simulated_2015}) of human land use.

@@

Additionally to the steady-state method \citep{calvo_buendia_ipcc_2019} and the detailed analysis of management data coming with it, SOC changes can be estimated using the IPCC Tier 1 approach of IPCC guidelines (\citep{eggleston_hs_ipcc_2006}, \citep{calvo_buendia_ipcc_2019}). Here, stocks are calculated via stock change factors (\(F^{\mathrm{SCF}}\)) given by the IPCC for the topsoil (0-30 cm) and based on observational data. Estimates of \(F^{\mathrm{SCF}}\) are differentiated by different crops, management and input systems (here summarized under \(m\)) reflecting different dynamics under changed in- and outflows without explicitly tracking these. Moreover, estimates of \(F^{\mathrm{SCF}}\) vary for different climatic zones (\(c\)) specified by the IPCC (see Fig. \ref{append:climatemap}). The actual \(SOC\) stocks are thus calculated based on a given reference stock \(SOC^{\mathrm{ref}}\) by

@@

with \(T_{c,i}\) being the translation matrix for grid cells \(i\) into corresponding climate zones \(c\). For this analysis, we use the default \(F^{\mathrm{SCF}}\) from the Tier 1 method of \citep{eggleston_hs_ipcc_2006} and \citep{calvo_buendia_ipcc_2019} as a comparison and consistency check for our more detailed Tier 2 steady-state approach.

@@

Land-use patterns are based on the Land-Use Harmonization 2 \citep{hurtt_harmonization_2020} data set, which we sum up from quarter-degree to half-degree resolution. We disaggregate the physical area of the five different cropland subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17 crop groups, applying the relative shares for each grid cell based on the country- and year-specific area harvested shares of FAOSTAT data \citep{faostat_faostat_2016} (see FAO2LUH2MAG\_croptypes.csv in \citep{mrcommons}). By calculating country-specific multicropping factors using FAOSTAT data, we are able to compute crop-group specific area harvested on grid cell level.
Land-use transitions are calculated as net area differences of the land-use data at half-degree resolution, considering no split up into crop-group specific areas but only total cropland and natural vegetation areas.

@@

Following the IPCC guidelines, we split the harvest index into a production (\(HI^{\mathrm{prod}}\)) and an area dependent (\(HI^{\mathrm{area}}\)) fraction \citep{eggleston_hs_ipcc_2006}. Note that deviating from \citep{bodirsky_n2o_2012} we use harvested instead of physical crop area to account for increased residue biomass due to multiple cropping and decreased residue amounts due to fallow land. We assume that all \(BGR\) are left in the soil, whereas \(AGR\) can be burned or harvested for other purposes such as feeding animals \citep{weindl_livestock_2017}, fuel or for material use.

A fixed share of the \(AGR\) is assumed to be burned on field depending on the per-capita income of the country. Following \citep{smil_nitrogen_1999}) we assume a burn share of \(25\%\) for low-income countries according to World Bank definitions (\(<\,1000\,\tfrac{USD}{yr\,cap}\)), 15\% for high-income (\(>\,10000\,\tfrac{USD}{yr\,cap}\)) and linearly interpolate shares for all middle-income countries depending on their per-capita income. Depending on the crop group, 80--90\% of the carbon in the crop residues burned in the fields is lost within the combustion process \citep{eggleston_hs_ipcc_2006}.

From our 17 crop groups, we compile four residue groups (straw, high- and low-lignin residues, residues without dual use), of which the first three are taken away from the field for other purposes (see mappingCrop2Residue.csv in \citep{mrcommons}). Residue feed demand for five different livestock groups is based on country-specific feed baskets (see \citep{weindl_livestock_2017}), that differentiate between the residue groups and take available \(AGR\) biomass as well as livestock productivity into account. We estimate a material-use share for the straw-residue group of 5\% and a fuel-share of 10\% for all used residue groups in low-income countries. For high-income countries, no withdrawal for material or fuel use is assumed, and use shares of middle-income countries are linearly interpolated based on per-capita income, following the same rationale as for the share of burnt residues described above. The remaining \(AGR\) as well as all \(BGR\) are expected to be left on the field. We limit high returning rates at \(10\unit{tC\,ha}^{-1}\) to correct for outliers.

To transform dry matter estimates into carbon and nitrogen, we compiled crop-group and plant-part specific carbon resp. nitrogen to dry matter (c/dm, n/dm) ratios (see \ref{append:Tablec2dm}).

@@

To highlight the impact of changing management effects and to assess the sensitivity of the model towards different initialization and parameter choices, we perform a set of scenario runs. In the following sections we outline name and idea of these scenarios (for technical implementation see \citep{mrsoil_karstens_2020}).

@@

As outlined in Sect. \citet{ref}(sec:initsoc) we assume landuse-type specific steady-state SOC stocks for the start year 1901 followed by a long spin-up period of 74 years ((\textit{Initial-lu})). As some SOC compartments decay over very long timescales, the initialisation setting might strongly effect the overall outcome of SOC stocks and changes. Thus we conduct counterfactual scenario \textit{Initial-natveg}, initializing the start year with the steady-state SOC under potential natural vegetation for all land-use types. We additionally combined the \textit{Initial-natveg} scenario with the (\textit{constManagement} scenario.

@@

Due to missing information about the litterfall composition, we assume woody potential natural vegetation for the lignin-to-carbon (short lg/c) and nitrogen-to-carbon (short n/c) ratios for all land under natural vegetation. This is however a strong simplification. Therefore, we test the sensitivity of the model for other natural vegetation parameterizations. Perennial grasses (scenario name \textit{LitterPNV-PerennialGrasses}) are given by the IPCC guidelines (\citep{calvo_buendia_ipcc_2019}), but for woody biomass we used the CENTURY configuration file (\citep{century_model_2000}) for trees as parameterization were not given. Whereas lg/c ratios ranges between \(0.09-0.35\) with an average of \(0.24\) (default value) for different tree compartments and different tree types, n/c ratios range between \(0.001-0.05\) with an average of \(0.015\) (default value). We decided to average over all tree compartments and tree types equally for the default values, since we have no data available on tree type distribution and tree compartment composition. We test however a range of different plant types as specified in Table \citet{ref}(tab:scenlitterpnv).

@@

\begin{figure}[h]
\includegraphics[width=18cm]{../ResultNotebooks/Output/Images/4panelfigure} \caption{(a): Distribution of total global SOC stocks for the first 30 cm on cropland shows high carbon stocks in high yielding areas. (b)+(c): Absolute (b) and relative (c) SOC stocks changes compared to a potential natural state showing different hot spots of SOC dynamics. Whereas absolute losses $\Delta SOC$ are often highest in temperate dry regions, relative losses $F^\mathrm{SCF}$ are often larger in tropical moist areas. (d): The SOC debt between SOC under historic land use and potential natural vegetation is decreasing over time, meaning net SOC gains on global croplands over the last decades. }\label{fig:SOCmaps}
\end{figure}

@@

Focusing of the temporal dimension of SOC stock changes (see Fig. \ref{fig:SOCmaps}(c)), global SOC debt has decreased around 15\% over the last couple of decades to \(22\unit{GtC}\). This leads to a sequestration rate of \(0.11\unit{GtC yr^{-1}}\) and considering our estimate of the global SOC stock of \(675\unit{GtC}\) in 2010 to \(0.2\) per mille per year.

\hypertarget{agricultural-management-effects-on-soc-debt}{%
\subsection{Agricultural management effects on SOC debt}\label{agricultural-management-effects-on-soc-debt}}

Figure \ref{fig:SOCscen}(a) reveals the relative impact of management effects by keeping carbon inflows from crop residue returning or manure application constant at the level of 1975 and neglecting the adoption of no-tillage practices. Without changes in management regimes, especially in C inflows from crop residues to the soil, global SOC debt on cropland would be increase at a rate of \(0.1\unit{GtC yr^{-1}}\). The constResidue scenario dominates the effect and is only single-off scenario that still losses SOC compared to natural vegetation at a rate of \(0.06\unit{GtC yr^{-1}}\). Whereas both the constManure and constTillage scenarios, show only small deviations from the historical values (sequestration rate of \(0.09\unit{GtC yr^{-1}}\) for both), the effect of no-tillage gets stronger till 2010.
The strong impact of almost doubling C inputs from crop residue biomass over a period of 35 years is visible in Fig. \ref{fig:SOCscen}(b) as well.

\begin{figure}[h]
\includegraphics[width=18cm]{../ResultNotebooks/Output/Images/scenario_horiz} \caption{(a) Global $\Delta SOC$ in GtC for different scenarios deviating from historical argicultural management by holding effects of carbon inflows from residues (constResidues), manure (constManure) constant or neglecting adoption of no-tillage practices (constTillage). ConstManagement combines all three modifications. Note that $\Delta SOC$ is defined as the difference of SOC under land-use compared to a natural vegetation state. Figure (b) shows the carbon inflows from crop residue and manure, underlining the strong impact of residues for SOC stock and SOC stock changes.}\label{fig:SOCscen}
\end{figure}

The results of the sensitivity analysis concerning the initialization of SOC stocks at the beginning of the spin-up phase are shown in Fig. \ref{fig:SOcinit}. The Initial-natveg scenario initializes the start year with the steady-state SOC under potential natural vegetation for all land-use types, compared to the default assumption (Initial-lu), which assumes landuse-type specific steady-state SOC stocks in 1901. The \(\Delta SOC\) estimation almost halves (from \textasciitilde26 GtC to \textasciitilde14 GtC) as well as the decrease in of \(\Delta SOC\) for over the last 35 years (from \textasciitilde4 GtC to \textasciitilde2.5 GtC) for the Inital-natveg scenario.

\begin{figure}[H]
\includegraphics[width=18cm]{../ResultNotebooks/Output/Images/scenario_init} \caption{Global $\Delta SOC$ for different SOC initialization choices in the start year 1901.}\label{fig:SOCinit}
\end{figure}

Figure \ref{fig:}

@@

\begin{figure}[h]
\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/CarbonBudget} \caption{Global carbon flows (at arrows) and SOC stocks (in boxes) within the agricultural system for the year 2010 (in MtC): Most important carbon sources on cropland are crop residues.}\label{fig:FlowFig}
\end{figure}

@@

Furthermore, \citep{sanderman_soil_2017} modeled historical trends based on agricultural land expansion without considering SOC variations due to time-variant agricultural management. \citep{pugh_simulated_2015} considered management effects like tillage and residue returning in a static way, but neither changes over time nor alignment to observed historical data like yields-levels or no-tillage areas were taken into account. The study moreover concludes that crop productivity gains (increasing yield levels by 18\%) do not lead to a~substantial decline in SOC debt (less than 1\% change).

@@

Modeling management effects at the global scale, however, comes with parametric and structural uncertainties. High SOC gains in the first 30 cm of the soil compared to natural vegetation in e.g.~the temperate croplands of Central Europe and UK, indicate suspicious high estimations of SOC inputs. As pointed out by \citep{keel_large_2017} and \citep{smith_how_2020}, carbon input calculations are highly sensitive to the choice of allometric functions determining below and above ground residue estimates from harvested quantities. \citep{keel_large_2017} question whether below ground residues might increase with a fixed root:shoot ratio rather than being independent of productivity gains. Moreover, the study pointed out that plant breeding shifts allometries, which might not be reflected in outdated data sources. Following this line of argument, SOC results calculated in this study may overestimate actual SOC stocks due to an overestimation in residue biomass, for which the ratio compared to the harvested organ is normally reduced due to breeding. However, with an evaluation of management effects (see Sect. \ref{sec:ipcccompare}), there is no indication for a systematic overrating of residue biomass. Rather, it is also likely that we are still missing carbon inputs e.g.~from cover- and inter-cropping to the soil.

@@

Generally, the limit to the first 30 cm of the soil profile follows the IPCC guidelines (\citep{eggleston_hs_ipcc_2006}, \citep{calvo_buendia_ipcc_2019}) and assumes, that most of the SOC dynamic happens in the topsoil. In this regard several aspects are over simplified within our approach. Firstly, distribution of carbon inputs into different soil layers are neglected and all carbon inputs are allocated to the topsoil. This particularly overestimates SOC stocks in the first 30 cm of soil below deeper rooting vegetation, which is certainly the case for most of the woody natural vegetated areas. Secondly changes to the subsoil due to tillage are neglected, which might be important to noticed, hence studies (\citep{powlson_limited_2014}) have shown, the subsoil to be a game changer in evaluating total SOC losses or gains for no-tillage systems. Thirdly, emissions from croplands on organic soils (like peat- and wetlands) are likely underestimate substantially.

@@

To validate our modeled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors (\citep{eggleston_hs_ipcc_2006}, \citep{calvo_buendia_ipcc_2019}) which are based on measurement data for croplands (see Table \ref{tab:SCFtable}). To allow for comparison, we aggregate our stock change factors to the four IPCC climate zones (Fig. \ref{append:climatemap}).

\textbackslash begin\{figure\}
\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/TableSCF_comparison} \textbackslash caption\{This table shows different estimates of stock change factors \(F_{SCF}\), defined as the SOC stock of managed croplands relative to the counterfactual SOC stock under natural vegetation. We compare average values for four IPCC climate zone classifications. IPCC2006 and IPCC2019 default factors (medium input) as well as altered factors for low and high input of organic matter without any other subsystem consideration (see Table 5.5 in \citep{eggleston_hs_ipcc_2006}, \citep{calvo_buendia_ipcc_2019} are compared to 1990 and 2010 results for historic input levels of this study.\}\label{fig:SCFtable}
\textbackslash end\{figure\}

Whereas our estimates are too low in the two tropical climate zones, temperate zone default factors are overestimated for the refinement of the IPCC guidelines (\citep{calvo_buendia_ipcc_2019}). Admittedly, the tropical factor differ substantially between the IPCC guidelines in 2006 (\citep{eggleston_hs_ipcc_2006}) and their refinement in 2019 (\citep{calvo_buendia_ipcc_2019}), leaving our tropical results in between.
Taking the oversimplified assumption, that tropical soils might suffer from insufficient C input rates (low) due to yield gaps , whereas temperate soils in developed regions might be highly managed and fertilized (high), the difference vanish partially. Our results show still especially in temperate dry regions -- the smallest region area-wise -- small deviations from natural SOC stocks. Considering the impact of irrigation ad fertilization on carbon-poor dryland soils, even factors above 1 (see Fig. \ref{fig:SOCmaps}(c)) might be not surprising.

>>>>

In Sect. \ref{sec:carbonbudget} we introduce the basic concept of SOC dynamics as applied in this study and explained in more detail within the refinement of the IPCC guidelines vol.~4 (\citep{calvo_buendia_ipcc_2019}). We additionally describe how we configured and extended the steady-state method (for model code see \citep{karstens_mrsoil_2020}). In Sect. \ref{sec:tier1} we shortly refer to the concept of stock change factors as outlined in the Tier 1 approach of the IPCC guidelines (\citep{ipcc_eggleston_2006}, \citep{calvo_buendia_ipcc_2019}).
Section \ref{sec:agrimanagement} provides a detailed description of the global, gridded management data used to drive the model, including crop production levels, residue input rates, manure amendments, and the adoption of irrigation and tillage practices (for model code see \citet{mrcommons_2020}). In Sect. \ref{sec:scenarios} we define the scenarios used to complement our historic model results.

@@

  \begin{table*}[h]
  \caption{Type and data sources for carbon inputs and parameterization to different land-use types }
  \begin{tabular}{p{0.15\textwidth} p{0.20\textwidth} p{0.25\textwidth} p{0.40\textwidth}}
  \tophline
  \textbf{land-use types}   & \textbf{source of carbon inputs} & \textbf{data source} & \textbf{nitrogen and lignin content} \\
  \middlehline
  \multirow{3}{*}{Cropland} & above ground residues & \multirow{3}{*}{\begin{minipage}[t]{1\columnwidth}\raggedright
                                                                      \cite{faostat_faostat_2016}, \\
                                                                      \cite{schaphoff_lpjml4_2018} (LPJmL),\\
                                                                      \cite{weindl_livestock_2017} \end{minipage}} & 
                                                      \multirow{3}{*}{\begin{minipage}[t]{1\columnwidth}\raggedright
                                                                      lg/c generic values based on Table 5.5B, 5.5C \\
                                                                      in \cite{calvo_buendia_ipcc_2019},\\
                                                                      crop-type specific n/c-ratios from
                                                                      \cite{bodirsky_mrcommons_2020}\end{minipage}} \\
                            & below ground residues &  &  \\ 
                            & manure                &  &  \\ 
                            \hline
  Natural vegetation        & annual litterfall &  \cite{schaphoff_lpjml4_2018} (LPJmL) 
                            & \begin{minipage}[t]{0.28\columnwidth}\raggedright\strut 
                                \cite{calvo_buendia_ipcc_2019} (CENTURY parameterization) \strut \end{minipage}\tabularnewline
 \bottomhline
 \end{tabular}
 \label{tab:datasourceinputs}
 \belowtable{}
 \end{table*}

@@

Note that the absolute changes \(\Delta SOC\) can be also interpreted as the SOC debt (\citep{sanderman_soil_2017}) due to human cropping activities; whereas relative changes \(F^{\mathrm{SCF}}\) can be considered stock change factors as defined within the IPCC guidelines of 2006 \citep{eggleston_ipcc_2006}. Moreover, \(\Delta SOC\) is equivalent to the negated cumulative SOC component of human land use change emission (\citep{pugh_simulated_2015}).

@@

Additionally to the steady-state method \citep{calvo_buendia_ipcc_2019} and the detailed analysis of management data coming with it, SOC changes can be estimated using the IPCC Tier 1 approach of IPCC guidelines (\citep{eggleston_ipcc_2006}, \citep{calvo_buendia_ipcc_2019}). Here, stocks are calculated via stock change factors (\(F^{\mathrm{SCF}}\)) given by the IPCC for the topsoil (0-30 cm) and based on observational data. Estimates of \(F^{\mathrm{SCF}}\) are differentiated by different crops, management and input systems (here summarized under \(m\)) reflecting different dynamics under changed in- and outflows without explicitly tracking these. Moreover, estimates of \(F^{\mathrm{SCF}}\) vary for different climatic zones (\(c\)) specified by the IPCC (see Fig. \ref{fig:CLIMzone}). The actual \(SOC\) stocks are thus calculated based on a given reference stock \(SOC^{\mathrm{ref}}\) by

@@

with \(T_{c,i}\) being the translation matrix for grid cells \(i\) into corresponding climate zones \(c\). For this analysis, we use the default \(F^{\mathrm{SCF}}\) from the Tier 1 method of \citep{eggleston_ipcc_2006} and \citep{calvo_buendia_ipcc_2019} as a comparison and consistency check for our more detailed Tier 2 steady-state approach.

@@

Land-use patterns are based on the Land-Use Harmonization 2 \citep{hurtt_harmonization_2020} data set, which we sum up from quarter-degree to half-degree resolution. We disaggregate the physical area of the five different cropland subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17 crop groups, applying the relative shares for each grid cell based on the country- and year-specific area harvested shares of FAOSTAT data \citep{faostat_faostat_2016} (see FAO2LUH2MAG\_croptypes.csv in \citep{bodirsky_mrcommons_2020}). By calculating country-specific multicropping factors using FAOSTAT data, we are able to compute crop-group specific area harvested on grid cell level.
Land-use transitions are calculated as net area differences of the land-use data at half-degree resolution, considering no split up into crop-group specific areas but only total cropland and natural vegetation areas.

@@

Following the IPCC guidelines, we split the harvest index into a production (\(HI^{\mathrm{prod}}\)) and an area dependent (\(HI^{\mathrm{area}}\)) fraction \citep{eggleston_ipcc_2006}. Note that deviating from \citep{bodirsky_n2o_2012} we use harvested instead of physical crop area to account for increased residue biomass due to multiple cropping and decreased residue amounts due to fallow land. We assume that all \(BGR\) are left in the soil, whereas \(AGR\) can be burned or harvested for other purposes such as feeding animals \citep{weindl_livestock_2017}, fuel or for material use.

A fixed share of the \(AGR\) is assumed to be burned on field depending on the per-capita income of the country. Following \citep{smil_nitrogen_1999}) we assume a burn share of \(25\%\) for low-income countries according to World Bank definitions (\(<\,1000\,\tfrac{USD}{yr\,cap}\)), 15\% for high-income (\(>\,10000\,\tfrac{USD}{yr\,cap}\)) and linearly interpolate shares for all middle-income countries depending on their per-capita income. Depending on the crop group, 80--90\% of the carbon in the crop residues burned in the fields is lost within the combustion process \citep{eggleston_ipcc_2006}.

From our 17 crop groups, we compile four residue groups (straw, high- and low-lignin residues, residues without dual use), of which the first three are taken away from the field for other purposes (see mappingCrop2Residue.csv in \citep{bodirsky_mrcommons_2020}). Residue feed demand for five different livestock groups is based on country-specific feed baskets (see \citep{weindl_livestock_2017}), that differentiate between the residue groups and take available \(AGR\) biomass as well as livestock productivity into account. We estimate a material-use share for the straw-residue group of 5\% and a fuel-share of 10\% for all used residue groups in low-income countries. For high-income countries, no withdrawal for material or fuel use is assumed, and use shares of middle-income countries are linearly interpolated based on per-capita income, following the same rationale as for the share of burnt residues described above. The remaining \(AGR\) as well as all \(BGR\) are expected to be left on the field. We limit high returning rates at \(10\unit{tC\,ha}^{-1}\) to correct for outliers.

To transform dry matter estimates into carbon and nitrogen, we compiled crop-group and plant-part specific carbon resp. nitrogen to dry matter (c/dm, n/dm) ratios (see \ref{tab:c2dm}).

@@

To highlight the impact of changing management effects and to assess the sensitivity of the model towards different initialization and parameter choices, we perform a set of scenario runs. In the following sections we outline name and idea of these scenarios (for technical implementation see \citep{karstens_mrsoil_2020}).

@@

As outlined in Sect. \ref{sec:initsoc} we assume landuse-type specific steady-state SOC stocks for the start year 1901 followed by a long spin-up period of 74 years ((\textit{Initial-lu})). As some SOC compartments decay over very long timescales, the initialisation setting might strongly effect the overall outcome of SOC stocks and changes. Thus we conduct counterfactual scenario \textit{Initial-natveg}, initializing the start year with the steady-state SOC under potential natural vegetation for all land-use types. We additionally combined the \textit{Initial-natveg} scenario with the (\textit{constManagement} scenario.

@@

Due to missing information about the litterfall composition, we assume woody potential natural vegetation for the lignin-to-carbon (short lg/c) and nitrogen-to-carbon (short n/c) ratios for all land under natural vegetation. This is however a strong simplification. Therefore, we test the sensitivity of the model for other natural vegetation parameterizations. Perennial grasses (scenario name \textit{LitterPNV-PerennialGrasses}) are given by the IPCC guidelines (\citep{calvo_buendia_ipcc_2019}), but for woody biomass we used the CENTURY configuration file (\citep{century_model_2000}) for trees as parameterization were not given. Whereas lg/c ratios ranges between \(0.09-0.35\) with an average of \(0.24\) (default value) for different tree compartments and different tree types, n/c ratios range between \(0.001-0.05\) with an average of \(0.015\) (default value). We decided to average over all tree compartments and tree types equally for the default values, since we have no data available on tree type distribution and tree compartment composition. We test however a range of different plant types as specified in Table \ref{tab:scenlitterpnv}.

@@

\begin{figure}[h]
\includegraphics[width=18cm]{../ResultNotebooks/Output/Images/4panelfigure} \caption{(a): Distribution of total global SOC stocks for the first 30 cm on cropland shows high carbon stocks in high yielding areas. (b)+(c): Absolute (b) and relative (c) SOC stocks changes compared to a potential natural state showing different hot spots of SOC dynamics. Whereas absolute losses $\Delta SOC$ are often highest in temperate dry regions, relative losses $F^\mathrm{SCF}$ are often larger in tropical moist areas. (d): $\Delta SOC$ between SOC under historic land use and potential natural vegetation is decreasing over time, meaning net SOC gains on global croplands over the last decades. }\label{fig:SOCmaps}
\end{figure}

@@

Focusing of the temporal dimension of SOC stock changes (see Fig. \ref{fig:SOCmaps}(c)), global \(\Delta SOC\) has decreased around 15\% over the last couple of decades to \(22\unit{GtC}\). This leads to a sequestration rate of \(0.11\unit{GtC yr^{-1}}\) and considering our estimate of the global SOC stock of \(675\unit{GtC}\) in 2010 to \(0.2\) per mille per year.

\hypertarget{agricultural-management-effects-on-delta-soc}{%
\subsection{\texorpdfstring{Agricultural management effects on \(\Delta SOC\)}{Agricultural management effects on \textbackslash Delta SOC}}\label{agricultural-management-effects-on-delta-soc}}

Figure \ref{fig:SOCscen}(a) reveals the relative impact of management effects by keeping carbon inflows from crop residue returning or manure application constant at the level of 1975 and neglecting the adoption of no-tillage practices. Without changes in management regimes, especially in C inflows from crop residues to the soil, global \(\Delta SOC\) on cropland would be increase at a rate of \(0.1\unit{GtC yr^{-1}}\). The constResidue scenario dominates the effect and is only single-off scenario that still losses SOC compared to natural vegetation at a rate of \(0.06\unit{GtC yr^{-1}}\). Whereas both the constManure and constTillage scenarios, show only small deviations from the historical values (sequestration rate of \(0.09\unit{GtC yr^{-1}}\) for both), the effect of no-tillage gets stronger till 2010.
The strong impact of almost doubling C inputs from crop residue biomass over a period of 35 years is visible in Fig. \ref{fig:SOCscen}(b) as well.

\begin{figure}
\includegraphics[width=14cm]{../ResultNotebooks/Output/Images/scenario_horiz} \caption{(a) Global $\Delta SOC$ in GtC for different scenarios deviating from historical argicultural management by holding effects of carbon inflows from residues (constResidues), manure (constManure) constant or neglecting adoption of no-tillage practices (constTillage). ConstManagement combines all three modifications. Note that $\Delta SOC$ is defined as the difference of SOC under land-use compared to a natural vegetation state. Figure (b) shows the carbon inflows from crop residue and manure, underlining the strong impact of residues for SOC stock and SOC stock changes.}\label{fig:SOCscen}
\end{figure}

The results of the sensitivity analysis concerning the initialization of SOC stocks at the beginning of the spin-up phase are shown in Fig. \ref{fig:SOCinit}. The Initial-natveg scenario initializes the start year with the steady-state SOC under potential natural vegetation for all land-use types, compared to the default assumption (Initial-lu), which assumes landuse-type specific steady-state SOC stocks in 1901. The \(\Delta SOC\) estimation almost halves (from \(~26\unit{GtC}\) to \(~14\unit{GtC}\)) as well as the decrease in of \(\Delta SOC\) for over the last 35 years (from \(~4\unit{GtC}\) to \(~2.5\unit{GtC}\)) for the Inital-natveg scenario.

\begin{figure}
\includegraphics[width=10cm]{../ResultNotebooks/Output/Images/scenario_init} \caption{Global $\Delta SOC$ for different SOC initialization choices in the start year 1901.}\label{fig:SOCinit}
\end{figure}

@@

\begin{figure}
\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/CarbonBudget} \caption{Global carbon flows (at arrows) and SOC stocks (in boxes) within the agricultural system for the year 2010 (in MtC): Most important carbon sources on cropland are crop residues.}\label{fig:FlowFig}
\end{figure}

@@

Furthermore, \citep{sanderman_soil_2017} modeled historical trends based on agricultural land expansion without considering SOC variations due to time-variant agricultural management. \citep{pugh_simulated_2015} considered management effects like tillage and residue returning in a static way, but neither changes over time nor alignment to observed historical data like yields-levels or no-tillage areas were taken into account. The study moreover concludes that crop productivity gains (increasing yield levels by 18\%) do not lead to a substantial decline in SOC debt (less than 1\% change).

@@

Modeling management effects at the global scale, however, comes with parametric and structural uncertainties. High SOC gains in the first 30 cm of the soil compared to natural vegetation in e.g.~the temperate croplands of Central Europe and UK, indicate suspicious high estimations of SOC inputs. As pointed out by \citep{keel_large_2017} and \citep{smith_how_2020}, carbon input calculations are highly sensitive to the choice of allometric functions determining below and above ground residue estimates from harvested quantities (see \ref{tab:c2dm} for coefficients used in this study). \citep{keel_large_2017} question whether below ground residues might increase with a fixed root:shoot ratio rather than being independent of productivity gains. Moreover, the study pointed out that plant breeding shifts allometries, which might not be reflected in outdated data sources. Following this line of argument, SOC results calculated in this study may overestimate actual SOC stocks due to an overestimation in residue biomass, for which the ratio compared to the harvested organ is normally reduced due to breeding. However, with an evaluation of management effects (see Sect. \ref{sec:ipcccompare}), there is no indication for a systematic overrating of residue biomass. Rather, it is also likely that we are still missing carbon inputs e.g.~from cover- and inter-cropping to the soil.

@@

Generally, the limit to the first 30 cm of the soil profile follows the IPCC guidelines (\citep{eggleston_ipcc_2006}, \citep{calvo_buendia_ipcc_2019}) and assumes, that most of the SOC dynamic happens in the topsoil. In this regard several aspects are over simplified within our approach. Firstly, distribution of carbon inputs into different soil layers are neglected and all carbon inputs are allocated to the topsoil. This particularly overestimates SOC stocks in the first 30 cm of soil below deeper rooting vegetation, which is certainly the case for most of the woody natural vegetated areas. Secondly changes to the subsoil due to tillage are neglected, which might be important to noticed, hence studies (\citep{powlson_limited_2014}) have shown, the subsoil to be a game changer in evaluating total SOC losses or gains for no-tillage systems. Thirdly, emissions from croplands on organic soils (like peat- and wetlands) are likely underestimate substantially.

@@

To validate our modeled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors (\citep{eggleston_ipcc_2006}, \citep{calvo_buendia_ipcc_2019}) which are based on measurement data for croplands (see Table \ref{tab:SCFglo}). To allow for comparison, we aggregate our stock change factors to the four IPCC climate zones (Fig. \ref{fig:CLIMzone}).

\% latex table generated in R 3.6.3 by xtable 1.8-4 package
\% Sun Dec 13 16:18:11 2020

\begin{table}[ht]
\centering
\begin{tabular}{rlllllll}
  \hline
 & Source & Input & Year & tropical moist & tropical dry & temperate dry & temperate moist \\ 
  \hline
1 & IPCC2006 & low & invariant & 0.44 & 0.55--0.61 & 0.74 & 0.66 \\ 
  2 & IPCC2006 & medium & invariant & 0.48 & 0.58--0.64 & 0.80 & 0.69 \\ 
  3 & IPCC2006 & high & invariant & 0.53 & 0.60--0.67 & 0.83 & 0.77 \\ 
  4 & IPCC2019 & low & invariant & 0.76 & 0.87 & 0.70--0.71 & 0.66--0.67 \\ 
  5 & IPCC2019 & medium & invariant & 0.83 & 0.92 & 0.76--0.77 & 0.69--0.70 \\ 
  6 & IPCC2019 & high & invariant & 0.92 & 0.96 & 0.79--0.80 & 0.77--0.78 \\ 
  7 & This Study & hist & 1990 & 0.57 & 0.61 & 0.78 & 0.76 \\ 
  8 & This Study & hist & 2010 & 0.64 & 0.68 & 0.83 & 0.83 \\ 
   \hline
\end{tabular}
\caption{$F^{\mathrm{SCF}}$ in comparison to IPCC Tier 1 default factors: Stock change factors are in good agreement with the default values of the IPCC in general. For the tropical regions the assumptions changed notablly from the guidelines in 2006 to the update in 2019. leaving our results in very good agreement with the old default assumptions. Default assumption are given under the assumption of medium input systems, which, considering the yield gap in mainly developing regions in the tropics, might be an overestimation and decrease $F^{\mathrm{SCF}}$ by additional 5-8 percent. Also modelled $F^{\mathrm{SCF}}$ have increased for all climates over time.} 
\end{table}

Whereas our estimates are too low in the two tropical climate zones, temperate zone default factors are overestimated for the refinement of the IPCC guidelines (\citep{calvo_buendia_ipcc_2019}). Admittedly, the tropical factor differ substantially between the IPCC guidelines in 2006 (\citep{eggleston_ipcc_2006}) and their refinement in 2019 (\citep{calvo_buendia_ipcc_2019}), leaving our tropical results in between.
Taking the oversimplified assumption, that tropical soils might suffer from insufficient C input rates (low) due to yield gaps , whereas temperate soils in developed regions might be highly managed and fertilized (high), the difference vanish partially. Our results show still especially in temperate dry regions -- the smallest region area-wise -- small deviations from natural SOC stocks. Considering the impact of irrigation ad fertilization on carbon-poor dryland soils, even factors above 1 (see Fig. \ref{fig:SOCmaps}(c)) might be not surprising.
